---
title: 'R Notebook - data from MSSE Weng et al., 2015'
author: "Michelle Voss"
date: "Nov 2018"
output:
  github_document:
    toc: true
    toc_depth: 1
    html_preview: false
---

**Citation**: Weng, T. B., Pierce, G. L., Darling, W. G., & Voss, M. W. (2015). Differential effects of acute exercise on distinct aspects of executive function. Medicine and science in sports and exercise, 47(7), 1460-1469.

**Code reviews for this notebook**: </br>
* James Kent, Tim Weng


# setup
Clear previous, load packages
```{r, include=FALSE}

rm(list=ls(all=TRUE))  #clear previous
  
library(car)
library(ggplot2)
library(knitr)
library(reshape2)
library(Hmisc)
library(psych)
library(afex)
library(tidyverse)  #load last, and dplyr comes with it
library(foreign)
library(jtools)
library(pander)
library(lsmeans)
```


## Load data from an spss file
Do not show in knit

```{r, include=FALSE}
data <- read.spss("acute_beh_analysis_50.sav", use.value.labels = TRUE, to.data.frame = TRUE)
attr(data, "variable.labels")
# make all the labels lowercase
names(data) <- tolower(names(data))
write.csv(data,"acute_beh_data_weng-etal-2015.csv", row.names = FALSE)
```




# demographics in Table 1
```{r}
demogs <- data %>% 
        select(age,gender,edu,height,weight,bmi,godin_score) 

pander(describe(demogs, skew=FALSE, ranges = FALSE, IQR = TRUE))
pander(describeBy(demogs, skew=FALSE, ranges = FALSE, IQR = TRUE, group = "gender"))
```




# paired t-tests for exercise conditions

```{r}
# get columns that match this pattern (| means "or")
exercise_vars <- grep(names(data), pattern = "^sub$|^exclude$|^countergroup$|active$|passive$|act$|pass$")
exercise_data <- select(data, exercise_vars)

# function for doing the ttest
desttest <- function(cols, data) {
    data[, cols] %>%
    describe(skew=FALSE, ranges = FALSE, IQR = TRUE) %>%
    pander()
  
  t.test(data[, cols[1]], 
         data[, cols[2]], 
         paired=TRUE, 
         conf.level=0.95)
}
```

## Exercise HR
```{r}
desttest(c("avghr_passive", "avghr_active"), exercise_data)
```

## Exercise %HRmax
```{r}
desttest(c("hrmax_pass", "hrmax_act"), exercise_data)
```


## RPE
```{r}
desttest(c("avgrpe_passive", "avgrpe_active"), exercise_data)

```



## FAS
```{r}
desttest(c("avgfas_passive", "avgfas_active"), exercise_data)
```


## FS
```{r}
desttest(c("avgfs_passive", "avgfs_active"), exercise_data)
```




# cognitive outcomes (n-back, flanker)
* percdiff=((post-pre)/pre)*100

## n-back
```{r}
nback_prepost <- data %>%
        filter(exclude == "include") %>%
        select(sub, countergroup, pre_passive_1b_acc:percdiff_active_2b_rt) %>%
        select(sub, countergroup, matches("pre|post")) %>%
        gather(key=condition, value=perf, -sub, -countergroup)

# now split condition column up to it's factors
nback_prepost <- separate(nback_prepost, "condition", c("time", "intensity", "load", "dvtype"), "_")

write.csv(nback_prepost, "acute_nback-prepost_weng-etal-2015.csv", row.names=FALSE)
```


```{r}
nback_percdiff <- data %>%
        filter(exclude == "include") %>%
        select(sub, countergroup, pre_passive_1b_acc:percdiff_active_2b_rt) %>%
        select(sub, countergroup, matches("percdiff")) %>%
        gather(key=condition, value=perf, -sub, -countergroup)

# now split condition column up to it's factors
nback_percdiff <- separate(nback_percdiff, "condition", c("drop", "intensity", "load", "dvtype"), "_")
nback_percdiff$drop <- NULL
write.csv(nback_percdiff, "acute_nback-percdiff_weng-etal-2015.csv", row.names=FALSE)
```


### set factors
```{r}
nback_prepost$time <- as.factor(nback_prepost$time)
nback_prepost$time <- relevel(nback_prepost$time, "pre")
nback_prepost$intensity <- as.factor(nback_prepost$intensity)
nback_prepost$intensity <- relevel(nback_prepost$intensity, "passive")
nback_prepost$load <- as.factor(nback_prepost$load)

nback_percdiff$intensity <- as.factor(nback_percdiff$intensity)
nback_percdiff$intensity <- relevel(nback_percdiff$intensity, "passive")
nback_percdiff$load <- as.factor(nback_percdiff$load)
```


### plot nback
```{r}
ggplot(subset(nback_prepost, dvtype == "acc"), aes(x=time, y=perf, group=sub)) + 
  geom_line() + 
  geom_point() +
  facet_grid(~intensity * load) +
  labs(title = "Accuracy", y = "Accuracy") + 
  theme(strip.text = element_text(face = "bold", size = 15, lineheight = 5.0), 
  strip.background = element_rect(colour = "black", size = 1))
```

```{r}
ggplot(subset(nback_prepost, dvtype == "rt"), aes(x = time, y = perf, group = sub)) + 
  geom_line() + 
  geom_point() +
  facet_grid(~intensity * load) +
  labs(title = "Reaction Time", y = "RT") + 
  theme(strip.text = element_text(face = "bold", size = 15, lineheight = 5.0), 
  strip.background = element_rect(colour = "black", size = 1))
```


```{r}
ggplot(subset(nback_percdiff,dvtype == "rt"), aes(x = intensity, y = perf, fill = load)) + 
  scale_fill_manual(values = c("#E3871C", "#F5BD78")) +
  geom_boxplot() + 
  labs(title = "Acute RT %change on N-back", y = "%Diff", x= "", fill = "load") +
  theme(title = element_text(size = 16, face = "bold"),
        axis.text.x = element_text(size = 16),
        axis.text.y = element_text(size = 15),
        axis.title.y = element_text(size = 16),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15))

ggplot(subset(nback_percdiff, dvtype == "rt"), aes(x = intensity, y = perf)) + 
  stat_summary(aes(y = perf), size = .5, fun.y = mean, geom = "bar", size = 1) + 
  stat_summary(fun.data = mean_cl_normal, geom = "errorbar", size = .5, width = .3) + 
        facet_grid(~load) +
        labs(title = "Acute RT %change on N-back", y = "%Diff", x = "") +
        theme(title = element_text(size = 16, face = "bold"),
        axis.text.x = element_text(size = 16),
        axis.text.y = element_text(size = 15),
        axis.title.y = element_text(size = 16),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15))

ggplot(subset(nback_percdiff, dvtype == "rt"), aes(x = load,y = perf)) + 
  stat_summary(aes(y = perf), size = .5, fun.y = mean, geom = "bar", size = 1) + 
  stat_summary(fun.data = mean_cl_normal, geom = "errorbar", size = .5, width = .3) + 
        facet_grid(~intensity) +
        labs(title = "Acute RT %change on N-back", y = "%Diff", x= "") +
        theme(title = element_text(size = 16, face = "bold"),
        axis.text.x = element_text(size = 16),
        axis.text.y = element_text(size = 15),
        axis.title.y = element_text(size = 16),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15))
```

```{r}
ggplot(subset(nback_percdiff,dvtype=="acc"), aes(x=intensity,y=perf,fill=load)) + 
  scale_fill_manual(values=c("#E3871C", "#F5BD78")) +
  geom_boxplot() + 
  labs(title="Acute ACC %change on N-back",y="%Diff",x="",fill="load") +
  theme(title=element_text(size=16, face='bold'),
        axis.text.x = element_text(size=16),
        axis.text.y = element_text(size=15),
        axis.title.y = element_text(size=16),
        legend.text = element_text(size=15),
        legend.title = element_text(size=15))

ggplot(subset(nback_percdiff,dvtype=="acc"), aes(x=intensity,y=perf)) + 
  stat_summary(aes(y = perf), size=.5, fun.y = mean, geom="bar",size=1) + 
  stat_summary(fun.data=mean_cl_normal, geom="errorbar", size=.5,width=.3) + 
        facet_grid(~load) +
        labs(title="Acute ACC %change on N-back",y="%Diff",x="") +
        theme(title=element_text(size=16, face='bold'),
        axis.text.x = element_text(size=16),
        axis.text.y = element_text(size=15),
        axis.title.y = element_text(size=16),
        legend.text = element_text(size=15),
        legend.title = element_text(size=15))

ggplot(subset(nback_percdiff,dvtype=="acc"), aes(x=load,y=perf)) + 
  stat_summary(aes(y = perf), size=.5, fun.y = mean, geom="bar",size=1) + 
  stat_summary(fun.data=mean_cl_normal, geom="errorbar", size=.5,width=.3) + 
        facet_grid(~intensity) +
        labs(title="Acute ACC %change on N-back",y="%Diff",x="") +
        theme(title=element_text(size=16, face='bold'),
        axis.text.x = element_text(size=16),
        axis.text.y = element_text(size=15),
        axis.title.y = element_text(size=16),
        legend.text = element_text(size=15),
        legend.title = element_text(size=15))
```


### nback stats

#### repeated measures anova with time as a factor
```{r}
aovACC <- aov_car(perf ~ time * intensity * load + Error(sub / time * intensity * load), subset(nback_prepost, dvtype == "acc"))
nice(aovACC)
```


#### repeated measures anova with percentdiff 

**reaction time**
```{r}
aovRT <- aov_car(perf ~ intensity * load + Error(sub / intensity * load), subset(nback_percdiff, dvtype == "rt"))
nice(aovRT)
aovRT_fitted <- lsmeans(aovRT, ~load|intensity)
aovRT_fitted
pairs(aovRT_fitted)
```


**accuracy**
```{r}
aovACC <- aov_car(perf ~ intensity * load + Error(sub / intensity * load), subset(nback_percdiff, dvtype == "acc"))
nice(aovACC)
aovACC_fitted <- lsmeans(aovACC, ~load|intensity)
aovACC_fitted
pairs(aovACC_fitted)
```

**Accuracy with glm and covariates **

Set to effect coding
```{r}
contrasts(nback_prepost$time) <- c(-.5, .5)
contrasts(nback_prepost$time)
contrasts(nback_prepost$intensity) <- c(-.5, .5)
contrasts(nback_prepost$intensity)
contrasts(nback_prepost$load) <- c(-.5, .5)
contrasts(nback_prepost$load)
```

```{r}
lm_acc <- lmer(perf ~ time * intensity * load + (1 + time + intensity + load | sub), data = subset(nback_prepost, dvtype == "acc"))
lm_acc2<-lmer(perf ~ time * intensity * load + (1 | sub), data = subset(nback_prepost, dvtype == "acc"))
anova(lm_acc, lm_acc2)
summary(lm_acc)
```

```{r}
cat_plot(lm_acc, pred = time, modx = intensity, mod2 = load, geom = "line")
```


## flanker
```{r}
flanker_prepost <- data %>%
        filter(exclude=="include") %>%
        select(sub,countergroup,pre_passive_inc_acc:percdiff_passive_con_rt) %>%
        select(sub,countergroup,matches("pre|post")) %>%
        gather(key=condition,value=perf,-sub,-countergroup)

# now split condition column up to it's factors
flanker_prepost <- separate(flanker_prepost, "condition", c("time", "intensity", "load", "dvtype"), "_")

write.csv(flanker_prepost, "acute_flanker-prepost_weng-etal-2015.csv", row.names=FALSE)
```



```{r}
flanker_percdiff <- data %>%
        filter(exclude=="include") %>%
        select(sub,countergroup,pre_passive_inc_acc:percdiff_passive_con_rt) %>%
        select(sub,countergroup,matches("percdiff")) %>%
        gather(key=condition,value=perf,-sub,-countergroup)

# now split condition column up to it's factors
flanker_percdiff <- separate(flanker_percdiff, "condition", c("drop", "intensity", "load", "dvtype"), "_")
flanker_percdiff$drop <- NULL

write.csv(flanker_percdiff,"acute_flanker-percdiff_weng-etal-2015.csv",row.names=FALSE)

```


### set factors
```{r}
flanker_prepost$time <- as.factor(flanker_prepost$time)
flanker_prepost$time <- relevel(flanker_prepost$time,"pre")
flanker_prepost$intensity <- as.factor(flanker_prepost$intensity)
flanker_prepost$intensity <- relevel(flanker_prepost$intensity,"passive")
flanker_prepost$load <- as.factor(flanker_prepost$load)

flanker_percdiff$intensity <- as.factor(flanker_percdiff$intensity)
flanker_percdiff$intensity <- relevel(flanker_percdiff$intensity,"passive")
flanker_percdiff$load <- as.factor(flanker_percdiff$load)
```



### plot flanker
```{r}
ggplot(subset(flanker_prepost,dvtype=="acc"), aes(x=time,y=perf,group=sub)) + 
  geom_line() + 
  geom_point() +
        facet_grid(~intensity*load) +
        labs(title="Accuracy",y="Accuracy") + 
        theme(strip.text = element_text(face="bold", size=15,lineheight=5.0), 
        strip.background = element_rect(colour="black", size=1))
```


```{r}
ggplot(subset(flanker_prepost,dvtype=="rt"), aes(x=time,y=perf,group=sub)) + 
  geom_line() + 
  geom_point() +
        facet_grid(~intensity*load) +
        labs(title="Reaction Time",y="RT") + 
        theme(strip.text = element_text(face="bold", size=15,lineheight=5.0), 
        strip.background = element_rect(colour="black", size=1))
```



```{r}
ggplot(subset(flanker_percdiff,dvtype=="rt"), aes(x=intensity,y=perf,fill=load)) + 
  scale_fill_manual(values=c("#E3871C", "#F5BD78","#CC9966")) +
  geom_boxplot() + 
  labs(title="Acute RT %change on flanker",y="%Diff",x="",fill="load") +
  theme(title=element_text(size=16, face='bold'),
        axis.text.x = element_text(size=16),
        axis.text.y = element_text(size=15),
        axis.title.y = element_text(size=16),
        legend.text = element_text(size=15),
        legend.title = element_text(size=15))

ggplot(subset(flanker_percdiff,dvtype=="rt"), aes(x=intensity,y=perf)) + 
  stat_summary(aes(y = perf), size=.5, fun.y = mean, geom="bar",size=1) + 
  stat_summary(fun.data=mean_cl_normal, geom="errorbar", size=.5,width=.3) + 
        facet_grid(~load) +
        labs(title="Acute RT %change on flanker",y="%Diff",x="") +
        theme(title=element_text(size=16, face='bold'),
        axis.text.x = element_text(size=16),
        axis.text.y = element_text(size=15),
        axis.title.y = element_text(size=16),
        legend.text = element_text(size=15),
        legend.title = element_text(size=15))

ggplot(subset(flanker_percdiff,dvtype=="rt"), aes(x=load,y=perf)) + 
  stat_summary(aes(y = perf), size=.5, fun.y = mean, geom="bar",size=1) + 
  stat_summary(fun.data=mean_cl_normal, geom="errorbar", size=.5,width=.3) + 
        facet_grid(~intensity) +
        labs(title="Acute RT %change on flanker",y="%Diff",x="") +
        theme(title=element_text(size=16, face='bold'),
        axis.text.x = element_text(size=16),
        axis.text.y = element_text(size=15),
        axis.title.y = element_text(size=16),
        legend.text = element_text(size=15),
        legend.title = element_text(size=15))
```


```{r}
ggplot(subset(flanker_percdiff,dvtype=="acc"), aes(x=intensity,y=perf,fill=load)) + 
  scale_fill_manual(values=c("#E3871C", "#F5BD78","#CC9966")) +
  geom_boxplot() + 
  labs(title="Acute ACC %change on flanker",y="%Diff",x="",fill="load") +
  theme(title=element_text(size=16, face='bold'),
        axis.text.x = element_text(size=16),
        axis.text.y = element_text(size=15),
        axis.title.y = element_text(size=16),
        legend.text = element_text(size=15),
        legend.title = element_text(size=15))

ggplot(subset(flanker_percdiff,dvtype=="acc"), aes(x=intensity,y=perf)) + 
  stat_summary(aes(y = perf), size=.5, fun.y = mean, geom="bar",size=1) + 
  stat_summary(fun.data=mean_cl_normal, geom="errorbar", size=.5,width=.3) + 
        facet_grid(~load) +
        labs(title="Acute ACC %change on flanker",y="%Diff",x="") +
        theme(title=element_text(size=16, face='bold'),
        axis.text.x = element_text(size=16),
        axis.text.y = element_text(size=15),
        axis.title.y = element_text(size=16),
        legend.text = element_text(size=15),
        legend.title = element_text(size=15))

ggplot(subset(flanker_percdiff,dvtype=="acc"), aes(x=load,y=perf)) + 
  stat_summary(aes(y = perf), size=.5, fun.y = mean, geom="bar",size=1) + 
  stat_summary(fun.data=mean_cl_normal, geom="errorbar", size=.5,width=.3) + 
        facet_grid(~intensity) +
        labs(title="Acute ACC %change on flanker",y="%Diff",x="") +
        theme(title=element_text(size=16, face='bold'),
        axis.text.x = element_text(size=16),
        axis.text.y = element_text(size=15),
        axis.title.y = element_text(size=16),
        legend.text = element_text(size=15),
        legend.title = element_text(size=15))
```

#### repeated measures anova with time as a factor
```{r}
aovACC<-aov_car(perf~time*intensity*load+Error(sub/time*intensity*load),subset(flanker_prepost,dvtype=="acc"))
nice(aovACC)
```


#### repeated measures anova with percentdiff 

**reaction time**
```{r}
aovRT<-aov_car(perf~intensity*load+Error(sub/intensity*load),subset(flanker_percdiff,dvtype=="rt"))
nice(aovRT)
aovRT_fitted<-lsmeans(aovRT,~load|intensity)
aovRT_fitted
pairs(aovRT_fitted)
```


**accuracy**
```{r}
aovACC<-aov_car(perf~intensity*load+Error(sub/intensity*load),subset(flanker_percdiff,dvtype=="acc"))
nice(aovACC)
aovACC_fitted<-lsmeans(aovACC,~load|intensity)
aovACC_fitted
pairs(aovACC_fitted)
```


**Accuracy with glm and covariates **

Set to effect coding
```{r}
contrasts(flanker_prepost$time) <- c(-.5,.5)
contrasts(flanker_prepost$time)
contrasts(flanker_prepost$intensity) <- c(-.5,.5)
contrasts(flanker_prepost$intensity)
flanker_prepost$load <- relevel(flanker_prepost$load,"neu")
# not sure if effect coding for 3-level variable is correct here
c<-contr.treatment(3)
my.coding<-matrix(rep(.25, 6), ncol=2)
my.simple<-c-my.coding
my.simple[my.simple==.75]<-.50
contrasts(flanker_prepost$load)<-my.simple
contrasts(flanker_prepost$load)
```


```{r}
lm_acc<-lmer(perf ~ time*intensity*load + (1 + time + intensity + load |sub), data=subset(flanker_prepost,dvtype=="acc"))
lm_acc2<-lmer(perf ~ time*intensity*load + (1 |sub), data=subset(flanker_prepost,dvtype=="acc"))
anova(lm_acc,lm_acc2)
summary(lm_acc)
```

```{r}
cat_plot(lm_acc, pred = time, modx = intensity, mod2 = load, geom="line")
```



```{r}
lm_rt<-lmer(perf ~ time*intensity*load + (1 + time + intensity + load |sub), data=subset(flanker_prepost,dvtype=="rt"))
lm_rt2<-lmer(perf ~ time*intensity*load + (1 |sub), data=subset(flanker_prepost,dvtype=="rt"))
anova(lm_rt,lm_rt2)
summary(lm_rt)
```


```{r}
cat_plot(lm_rt, pred = time, modx = intensity, mod2 = load, geom="line")
```